### 2025-09-20 YEAR2/PHY293_LectureNotes/main.tex
sept20

\end{example}
\section{The Wave Equation}

>>>>

\end{example}
\subsection{Normal Modes}
\paragraph{Oscilating Masses Connected by Springs} Consider two masses \( m_1 = m_2 = m \) connected by three spring with spring constant \( k \) and fixed to walls on either side. The equations of motion for the two masses are given by:
\begin{equation}
    \begin{cases}
        m\ddot{x}_1 &= -k x_1 + k (x_2 - x_1) \quad (a) = m\ddot{x}_1 = -2k x_1 + k x_2 \\
        m\ddot{x}_2 &= -k x_2 - k (x_2 - x_1) \quad (b) = m\ddot{x}_2 = k x_1 - 2k x_2
    \end{cases}
\end{equation}
To solve these equations, we assume solutions of the form:
$$
    \begin{cases}
        x_1(t) &= A \cos(\omega t + \phi) \\
        x_2(t) &= B \cos(\omega t + \phi)
    \end{cases}
$$
Substituting these assumed solutions into the equations of motion, we get:
\begin{equation} \label{eq:coupled_springs}
    \begin{cases}
        \frac{A}{B} &= \frac{k}{m\omega^2 - 2k} \\
        \frac{A}{B} &= \frac{m\omega^2 - 2k}{k}
    \end{cases}
\end{equation}
We can deduce that $A = \pm B$  
We can set the equation:
$$
    (m\omega^2 - 2k)^2 = k^2 \implies m\omega^2 - 2k = \pm k
$$

So we can derive two normal mode frequencies:
\begin{equation}
    \begin{cases}
        \omega_1 &= \sqrt{\frac{k}{m}} \quad \text{(in-phase mode)} \\
        \omega_2 &= \sqrt{\frac{3k}{m}} \quad \text{(out-of-phase mode)}
    \end{cases}
\end{equation}

\paragraph{Solving as an Eigenproblem} Alternatively, we can create a system of equations from \eqref{eq:coupled_springs}:
\begin{equation}
    \begin{bmatrix}
        \frac{2k}{m} & -\frac{k}{m} \\
        -\frac{k}{m} & \frac{2k}{m}
    \end{bmatrix}
    \begin{bmatrix}
        A \\ B
    \end{bmatrix}
    = \omega^2
    \begin{bmatrix}
        A \\ B
    \end{bmatrix}
\end{equation}
To find the eigenvalues \( \omega^2 \), we solve the characteristic equation:
$$
    \text{det}\left(\begin{bmatrix}
        \frac{2k}{m} - \omega^2 & -\frac{k}{m} \\
        -\frac{k}{m} & \frac{2k}{m} - \omega^2
    \end{bmatrix}\right) = 0
$$
Calculating the determinant, we have:
$$
    \left(\frac{2k}{m} - \omega^2\right)^2 - \left(-\frac{k}{m}\right)^2 = 0
$$
Expanding and simplifying, we get:
$$
    \omega^4 - \frac{4k}{m}\omega^2 + \frac{5k^2}{m^2} = 0
$$
This gives us the two normal mode frequencies:
\begin{equation}
    \begin{cases}
        \omega_1^2 &= \frac{k}{m} \implies \omega_1 = \sqrt{\frac{k}{m}} \\
        \omega_2^2 &= \frac{3k}{m} \implies \omega_2 = \sqrt{\frac{3k}{m}}
    \end{cases}
\end{equation}

\begin{example}
    CConsider two equal masses $m$ suspended from identical springs of spring constant $k$. The masses are hanged on the celing. We can solve this system's frequency via the eigenvalue method. The equations of motion are:
    $$
        \begin{cases}
            m\ddot{x}_2 &= -k(x_2 - x_1)\\
            m\ddot{x}_1 &= -k x_1 + k(x_2 - x_1) = -2k x_1 + k x_2
        \end{cases}
    $$
    Assume solutions of the form:
    $$        
    \begin{cases}
            x_1(t) &= A \cos(\omega t + \phi) \\
            x_2(t) &= B \cos(\omega t + \phi)
        \end{cases}
    $$
    Substituting these assumed solutions into the equations of motion, we get:
    $$
        \begin{cases}
            -A \omega^2 &= -\frac{k}{m}(-2A + B) \\
            -B \omega^2 &= -\frac{k}{m}(A - B)
        \end{cases}
    $$
    We can rewrite this as an eigenvalue problem:
    $$
        \begin{bmatrix}
            \frac{2k}{m} & -\frac{k}{m} \\
            -\frac{k}{m} & \frac{k}{m}
        \end{bmatrix}
        \begin{bmatrix}
            A \\ B
        \end{bmatrix}
        = \omega^2
        \begin{bmatrix}
            A \\ B
        \end{bmatrix}
    $$
    To find the eigenvalues \( \omega^2 \), we solve the characteristic equation:
    $$
        \text{det}\left(\begin{bmatrix}
            \frac{2k}{m} - \omega^2 & -\frac{k}{m} \\
            -\frac{k}{m} & \frac{k}{m} - \omega^2
        \end{bmatrix}\right) = 0
    $$
    Calculating the determinant, we have:
    $$
        \omega^2 = \frac{k}{2m}(3\pm \sqrt{5})
    $$
\end{example}

\section{The Wave Equation}
